---
title: "Retraining the ubi-site score"
output: html_document
---

# Background
This shows how to retrain the model using the functions in `scripts/model_training.R`,
`scripts/feature_processing.R`, and `scripts/pipeline.R`.

# Initialise
```{r}
source("scripts/model_training.R")
source("scripts/feature_processing.R")
source("scripts/pipeline.R")
```

# Load data
`datFeats` holds the processed features for every candidate site (rownames
`"<uniprot>_<site>"`), built from the raw per-lysine features in `allLys_HsRevi_feats.csv`
by `prepareFeatures()`. `trainSites` is a list of the form `list(pos = ..., neg = ...)`,
giving the rownames of `datFeats` to use as the positive/negative training sites, built by
`loadDefaultTrainSites()` from the sites originally used to train the model.
```{r}
datFeats <- prepareFeatures("data")
trainSites <- loadDefaultTrainSites("data")
```

# Example 1: retrain with all the defaults
```{r}
resLst <- runTrainingPipeline(
  datFeats = datFeats,
  trainSites = trainSites
)

resLst$AUCdf          # AUC of each train-test split, on its own held-out test set
resLst$modsPredsTest_df  # the underlying test-set predictions behind AUCdf
resLst$allPredsDf     # median predicted score per site in datFeats, across all splits
```

# Example 2: custom training sites and repeated splits
Supply your own positive/negative training sites (as rownames of `datFeats`), and generate
5 repeats of the 5-fold split instead of 1.
```{r}
myTrainSites <- list(
  pos = trainSites$pos,
  neg = trainSites$neg[1:5000]
)

resLst_custom <- runTrainingPipeline(
  datFeats = datFeats,
  trainSites = myTrainSites,
  numReps = 5
)
```

# Example 3: prepareFeatures() with custom sites
`sites` can be any vector of `uniprot_site` values representing lysines in the reviewed human proteome, not just the ones already used for
training - e.g. adding a couple of extra sites of interest alongside `trainSites`:
```{r}
datFeats_extraSites <- prepareFeatures(
  "data",
  sites = c("Q9NVI1_K523", "Q9BXW9_K561", trainSites$pos, trainSites$neg)
)
```

# Example 4: supplying your own train-test splits
If you'd rather build the train-test splits yourself (e.g. a different splitting
strategy), build a training set with `buildTrainingData()`, construct a list of
`list(trainDat = ..., testDat = ...)` splits from it, and pass that in directly.
```{r}
trainDat <- buildTrainingData(datFeats = datFeats, posSites = myTrainSites$pos, negSites = myTrainSites$neg)
myCustomSplits <- makeFoldsProtstrat(seed = 1, k = 5, data = trainDat)  # or your own splitting logic

resLst_customSplits <- runTrainingPipeline(
  datFeats = datFeats,
  trainSites = myTrainSites,
  customSplits = myCustomSplits
)
```

# Example 5: customising feature processing
`prepareFeatures()` runs: subset to `sites` (default: `loadDefaultPredSites()`) -> subset to
`features` -> mean-impute `meanImputeFeats` (default `FinalBoxProcNum`) -> drop sites with
any remaining missing feature values -> log-transform `logFeats` -> center and scale.

Mean-impute an extra feature:
```{r}
datFeats_customImpute <- prepareFeatures("data", meanImputeFeats = c("FinalBoxProcNum", "NSP_rsa"))
```

Restrict processing to a chosen set of sites, e.g. only the sites used for training:
```{r}
datFeats_subset <- prepareFeatures("data", sites = c(trainSites$pos, trainSites$neg))
```

Select only a subset of features:
```{r}
datFeats_subsetFeats <- prepareFeatures("data", features = c("FinalBoxProcNum", "NSP_rsa", "PRIDE_gold"))
```

Skip processing entirely and get the raw features (still subsetted to `features`/`sites`):
```{r}
datFeats_raw <- prepareFeatures("data", raw = TRUE)
```